### 2023-01-11 definitions.tex
Remrking that Ozkan and all have put down a condition for how to ensure a basis of up post*(s) is computable. The goal with a proof not using such a basis would then be to have a lesser condition.

\newcommand{\pred}{\textsc{Pred}}
\renewcommand{\succ}{\textsc{Succ}}

\noindent
The set of {\em successors} of a configuration $s \in S$ is deﬁned as
$\succ(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s \xrightarrow{a} s'\}$.
A configuration without successors is called a {\em dead end}.  \\
The set of {\em predecessor} of a configuration $s \in S$ is deﬁned as
$\pred(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s' \xrightarrow{a} s\}$.

@@

We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure). We use similar notation for $\succ$ and $\pred$...

@@

A transition system is {\em finitely branching} if all $\succ(s)$ are finite. 
\textcolor{blue}{We restrict our attention to finitely branching TSs.}

@@

If $S$ is a WSTS with strong monotony and $I \subseteq S$ is an upward-closed set, the sets $\pred(I )$, and $\pred^{\geq k}(I )$ for
every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(I)$.
Remark that $I^{k+1} = I \cup \pred(I^k)$, 
%thus $I^{k+1} = I^k$ implies $k(I) \leq k$.
which ensures the following.

@@

\begin{proof}
We have to show that we can compute a basis of $I^{k+1}$ if we are given a basis of $I^k $. 
Then the
decidability of the stop condition follows directly. Let $B$ be a basis of $I^k$. 
We have
$$I^{k+1} = I \cup \pred(I^k ) = I \cup
\bigcup_{s' \in B}
\pred(\uparrow \{s' \}).$$
\alain{n'est-ce pas $I^{k+1} = I^k \cup \pred(I^k ) = I^k \cup
\bigcup_{s' \in B}
\pred(\uparrow \{s' \}).$ ?}

@@

\begin{proof}
From Fact~\ref{stop condition}, there exists $k \in \N$ such that
$\pred^*(I) = I^{k_m}$. We can compute this $k_m$ by iteratively computing
$I^{k+1}$ from $I^k$, checking $I^{k+1}=I^k$, 
returning $k_m$ if that is the case.
Then, because $k_m$-resilience is decidable, 
checking $\uparrow post^*(s) \cap J \subseteq I^k = \pred^*(I)$ is,
and resilience is decidable.
% \textcolor{red}{à faire Mathieu}
\end{proof}

However, the hypothesis that $\uparrow post^*$ is computable cannot be tested in the WSTS framework.

\begin{proposition}
There exist classes of WSTSs with strong upward compatibility for which there don't exist an algorithm computing a basis of $\uparrow post^*$.
\end{proposition}

\begin{proof}
Let us show that Reset VASS and LCS don't enjoy the property that $\uparrow post^*$ is computable.
Suppose that one are able to compute a finite basis of $\uparrow post^*$ for Reset VASS. Then, on ewould be able to decide whether $0$ is reachable in counter $i$ by examining if there is some vector $v$ in the basis such that $v(i)=0$. But reachability of $0$ in a counter $i$ is undecidable for Reset VASS. sont WSTS with strong compatibility. et l'accessibilité est undecidable.
$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
\textcolor{red}{à relire par Mathieu}
\end{proof}

>>>>

\newcommand{\pred}{\textsf{pred}}
\newcommand{\post}{\textsf{post}}
% \renewcommand{\succ}{\textsf{Succ}}

\noindent
The set of {\em successors} of a configuration $s \in S$ is deﬁned as
$\post(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s \xrightarrow{a} s'\}$.
A configuration without successors is called a {\em dead end}.  \\
The set of {\em predecessor} of a configuration $s \in S$ is deﬁned as
$\pred(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s' \xrightarrow{a} s\}$.

@@

We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure). We use similar notation for $\post$ and $\pred$...

@@

A transition system is {\em finitely branching} if all $\post(s)$ are finite. 
\textcolor{blue}{We restrict our attention to finitely branching TSs.}

@@

If $S$ is a WSTS with strong monotony and $I \subseteq S$ is an upward-closed set, the sets $\pred(I )$, and $\pred^{\geq k}(I )$ for
every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(I)$.

\begin{remark}
$I^{k+1} $ can be rewritten $I^{k+1}= \bigcup_{0 \leq j \leq k+1} \pred^j(I) = 
I \cup \bigcup_{1 \leq j \leq k+1} \pred^j(I) =
I \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(I))
=  I \cup \pred(I^k)$.
\end{remark}

This ensures the following.

@@

\begin{proof}
We have to show that we can compute a basis of $I^{k+1}$ if we are given a basis of $I^k $. 
Then the
decidability of the stop condition follows directly. Let $B$ be a basis of $I^k$. 
We have
$$I^{k+1} = I \cup \pred(I^k ) = I \cup
\bigcup_{s' \in B}
\pred(\uparrow \{s' \}).$$
\alain{n'est-ce pas $I^{k+1} = I^k \cup \pred(I^k ) = I^k \cup
\bigcup_{s' \in B}
\pred(\uparrow \{s' \}).$ ?}
\mathieu{À strictement parler, c'est  $I^{k+1} = \bigcup_{0 \leq j \leq k+1} \pred^j(I) = 
I \cup \bigcup_{1 \leq j \leq k+1} \pred^j(I) =
I \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(I))
=  I \cup \pred(I^k)$. Je vais rajouter une remarque au dessus.}

@@

\begin{proof}
From Fact~\ref{stop condition}, there exists $k \in \N$ such that
$\post^*(I) = I^{k_m}$. We can compute this $k_m$ by iteratively computing
$I^{k+1}$ from $I^k$, checking $I^{k+1}=I^k$, 
returning $k_m$ if that is the case.
Then, because $k_m$-resilience is decidable, 
checking $\uparrow post^*(s) \cap J \subseteq I^k = \pred^*(I)$ is,
and resilience is decidable.
% \textcolor{red}{à faire Mathieu}
\end{proof}

\"Ozkan goes on to prove that for finite-branching WSTS with strong monotonicity, a basis of $\uparrow \post^*(s)$ is computable for every given state $s$ iff the downward-closed problem is decidable, i.e. given a state $s$ of a WSTS
% in the regarded class 
with strong monotonicity 
and a decidable downward-closed set $J$, it can be decided whether $\exists s' \in J ~ s \to^* s'$ (Proposition~1 in \cite{DBLP:conf/gg/Ozkan22}). The characterization is used to show that Petri nets are $\post^*$ -effective. It is well-known that Petri nets constitute WSTS with storng monotonicity (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS0}).
However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the WSTS framework.

\begin{proposition}
There exist classes of WSTSs with strong upward compatibility for which there don't exist an algorithm computing a basis of $\uparrow \post^*$.
\end{proposition}

\begin{proof}
Let us show that Reset VASS and LCS don't enjoy the property that $\uparrow \post^*$ is computable.
Suppose that one are able to compute a finite basis of $\uparrow post^*$ for Reset VASS. Then, on ewould be able to decide whether $0$ is reachable in counter $i$ by examining if there is some vector $v$ in the basis such that $v(i)=0$. But reachability of $0$ in a counter $i$ is undecidable for Reset VASS. sont WSTS with strong compatibility. et l'accessibilité est undecidable.
$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
\textcolor{red}{à relire par Mathieu}
\end{proof}
